---
title: "Reproducible Research Course Project Week2"
author: "Yue Li"
date: "January 29, 2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background  

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.  


This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Data Loading and Preparation  

First, data is downloaded from the Forked repository as a zip file:

```{r loading}
setwd("~/Desktop/Coursera All/Rep Data")
mydata <- read.csv(unz("activity.zip", "activity.csv"))
dim(mydata)
head(mydata)
summary(mydata)
```

From the result above, we can see there are 17568 records and 3 variables in the data. For the variable **steps**, there are 2304 records have NA values. In the data processing step, we will add a new variable indicating the weekdays, and create a cleaned dataset without NAs:

```{r processing}
mydata$day <- mydata$day <- weekdays(as.Date(mydata$date))
mydata_clean <- mydata[!is.na(mydata$steps),]
```

## What is mean total number of steps taken per day?  

1. Calculate the total number of steps taken per day

```{r question1}
steps_day <- aggregate(steps ~ date, data=mydata_clean, FUN=sum)
head(steps_day)

```

2. Make a histogram of the total number of steps taken each day

```{r hist}
hist(steps_day$steps, xlab="Total Steps per Day", ylab="Frequency", main = "Total number of steps taken per day",breaks=seq(0,25000,by=2500),ylim=c(0,20),col="lightblue")
```

3.Calculate and report the mean and median of the total number of steps taken per day

```{r}
#Mean
mean(steps_day$steps)

#Median
median(steps_day$steps)
```

## What is the average daily activity pattern?  

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r timeplot}
steps_interval <- aggregate(steps ~ interval, data = mydata_clean, FUN = mean) 
plot(steps_interval, type = "l")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  

```{r}
#maximum steps
maxsteps <- max(steps_interval$steps)
#the interval contains the maximum average number of steps
steps_interval[steps_interval$steps==maxsteps,1]
```

## Imputing missing values  

1. Calculate and report the total number of missing values in the dataset 
```{r}
sum(is.na(mydata))
```

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated  

My strategy for filling in NAs will be to substitute the missing steps with the mean steps of the respective 5-minute interval

```{r}
imputed_steps <- steps_interval$steps[match(mydata$interval, steps_interval$interval)]
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
mydata_imputed <- transform(mydata, steps = ifelse(is.na(mydata$steps), yes = imputed_steps, no = mydata$steps))
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
steps_day_impute <- aggregate(steps ~ date, data=mydata_imputed, FUN=sum)
hist(steps_day_impute$steps, xlab="Total Steps per Day", ylab="Frequency", main = "Total number of steps taken per day",breaks=seq(0,25000,by=2500),col="lightblue")

#Mean
mean(steps_day_impute$steps)

#Median
median(steps_day_impute$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day  

```{r}
## create new category based on the days of the week
mydata_imputed$DayCategory <- ifelse(mydata_imputed$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
```

2.Make a panel plot containing a time series plot (i.e. 𝚝𝚢𝚙𝚎 = "𝚕") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
library(ggplot2)
steps_daycat <- aggregate(steps ~ interval + DayCategory, mydata_imputed, FUN=mean)
plot <- ggplot(steps_daycat, aes(x = interval , y = steps, color = DayCategory)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~DayCategory, ncol = 1, nrow=2)
plot
```

